---
title: "8_Samples_ED_and_BR_filtration"
author: "Maria"
date: "2024-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r libraries, echo=FALSE}
library(dplyr)
library(Matrix)
library(tibble)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(DropletUtils)
library(glue)

```
##### Basic functions
*`UMI_per_CB`* returns a data frame for a single sample, containing the UMI counts per cell barcode (CB) and the ordinal number of the CB in a sorted array.

```{R, UMI_per_CB for 1 sample}

# возвращает DF для одного образца, с UMI_CB и порядковым номером CB в отсортированном массиве
UMI_per_CB <- function(counts_path_s1, barcodes_path_s1, genes_path_s1, counts_path_s2, barcodes_path_s2, genes_path_s2){
  
  #Read files
  counts <- Matrix::readMM(counts_path_s1)
  genes <- readr::read_tsv(genes_path_s1, col_names = FALSE)
  gene_ids <- genes$X1
  cell_ids <- readr::read_tsv(barcodes_path_s1, col_names = FALSE, show_col_types = FALSE)$X1
  
  #Assign row and colnames
  rownames(counts) <- gene_ids
  colnames(counts) <- cell_ids
  
  #Calculate UMI per cell barcode distribution
  umi_per_CB <- Matrix::colSums(counts)[Matrix::colSums(counts) > 0]
  
  #Calculate gene per cell barcode distribution
  gene_per_CB <- Matrix::colSums(counts > 0)[Matrix::colSums(counts) > 0]
  
  
  #Create sorted UMI_count DF
  DF_umi_per_CB <- as.data.frame(umi_per_CB) %>%
    arrange(desc(umi_per_CB)) %>%
    mutate(cb = as.numeric(rownames(as.data.frame(umi_per_CB))))
  return (DF_umi_per_CB)
}
```
*filter_barcodes_with_emptyDrops* performs filtration with EmptyDrops
*return_filtered_barcodes_or_indeces* return list of selected barcodes or barcodes indeces

```{R, ED filtratin function}
#  functions for  Empty Drops filtration and function for gettting list of filtered barcodes

filter_barcodes_with_emptyDrops <- function(sparse_matrix, lower = 100, test.ambient = TRUE) {
  emptyDrops_df = emptyDrops(sparse_matrix, lower = lower, test.ambient = test.ambient)
  emptyDrops_df$FDR[is.na(emptyDrops_df$FDR)] <- 1 # выставляю всем NA FDR единичку
  return(emptyDrops_df)
}

return_filtered_barcodes_or_indeces <- function(emptyDrops_df, fdr_threshold = 0.05, return_indeces = FALSE) {
  filtered_barcodes_indeces <- which(emptyDrops_df$FDR < fdr_threshold) #список со всеми отфильтр. баркодами
  
  if (return_indeces) {
    return(filtered_barcodes_indeces)
  }
  return(rownames(emptyDrops_df)[filtered_barcodes_indeces])
}

```

# Paths to files
```{R paths, echo=FALSE, warnings = FALSE}

counts_path_s1 <- "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_noPEG_R1/Solo.out/GeneFull/raw/matrix.mtx"
barcodes_path_s1 <- "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_noPEG_R1/Solo.out/GeneFull/raw/barcodes.tsv"  
genes_path_s1 <- "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_noPEG_R1/Solo.out/GeneFull/raw/features.tsv"
  
counts_path_s2 <- "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_noPEG_R2/Solo.out/GeneFull/raw/matrix.mtx"
barcodes_path_s2 <- "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_noPEG_R2/Solo.out/GeneFull/raw/barcodes.tsv "
genes_path_s2 <- "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_noPEG_R2/Solo.out/GeneFull/raw/features.tsv"

counts_path_s3 <- "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_PEG_R1/Solo.out/GeneFull/raw/matrix.mtx"
barcodes_path_s3 <- "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_PEG_R1/Solo.out/GeneFull/raw/barcodes.tsv" 
genes_path_s3 <-   "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_PEG_R1/Solo.out/GeneFull/raw/features.tsv"
  
counts_path_s4 <- "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_PEG_R2/Solo.out/GeneFull/raw/matrix.mtx"
barcodes_path_s4 <- "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_PEG_R2/Solo.out/GeneFull/raw/barcodes.tsv " 
genes_path_s4 <- "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_PEG_R2/Solo.out/GeneFull/raw/features.tsv "

counts_path_s5 <- "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_noPEG_R1/Solo.out/GeneFull/raw/matrix.mtx"
barcodes_path_s5 <- "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_noPEG_R1/Solo.out/GeneFull/raw/barcodes.tsv" 
genes_path_s5 <- "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_noPEG_R1/Solo.out/GeneFull/raw/features.tsv"

counts_path_s6 <- "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_noPEG_R2/Solo.out/GeneFull/raw/matrix.mtx"
barcodes_path_s6 <- "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_noPEG_R2/Solo.out/GeneFull/raw/barcodes.tsv"  
genes_path_s6 <- "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_noPEG_R2/Solo.out/GeneFull/raw/features.tsv"

counts_path_s7 <- "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_PEG_R1/Solo.out/GeneFull/raw/matrix.mtx"
barcodes_path_s7 <- "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_PEG_R1/Solo.out/GeneFull/raw/barcodes.tsv" 
genes_path_s7 <-   "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_PEG_R1/Solo.out/GeneFull/raw/features.tsv" 

counts_path_s8 <- "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_PEG_R2/Solo.out/GeneFull/raw/matrix.mtx"
barcodes_path_s8 <- "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_PEG_R2/Solo.out/GeneFull/raw/barcodes.tsv" 
genes_path_s8 <- "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_PEG_R2/Solo.out/GeneFull/raw/features.tsv"

counts_paths <- c(counts_path_s1, counts_path_s2, counts_path_s3, 
                  counts_path_s4, counts_path_s5, counts_path_s6,
                  counts_path_s7, counts_path_s8)

barcodes_paths <- c(barcodes_path_s1, barcodes_path_s2, barcodes_path_s3, 
                    barcodes_path_s4, barcodes_path_s5, barcodes_path_s6,
                    barcodes_path_s7, barcodes_path_s8)

genes_paths <- c(genes_path_s1, genes_path_s2, genes_path_s3, 
                  genes_path_s4, genes_path_s5, genes_path_s6,
                  genes_path_s7, genes_path_s8)
  
```
##### Create df with UMIs_per_CB for each sample

```{R sample matrix, warning = FALSE, message = FALSE}

s <- list()

for (i in 1:length(counts_paths)) {
  var_name <- paste0("s", i)
  s[[i]] <- UMI_per_CB(counts_paths[i], barcodes_paths[i], genes_paths[i])
  assign(var_name, s[[i]])
}
```

Get list of filtered barcodes for  each sample using EmptyDrops for each sample
```{R, ED filtration}
matrix_list <- list()
for (i in 1:length(counts_paths)) {
  # Read the matrix
  matrix_raw <- readMM(counts_paths[i])
  cell_ids <- readr::read_tsv(barcodes_paths[i], col_names = FALSE, show_col_types = FALSE)$X1
  colnames(matrix_raw) <- cell_ids
  matrix_list[[i]] <- matrix_raw
}

ED_CB_list <- list()

for (i in 1:length(matrix_list)) {
  emptyDrops_df <- filter_barcodes_with_emptyDrops(matrix_list[[i]])
  ED_CB_list[[i]] <- return_filtered_barcodes_or_indeces(emptyDrops_df, fdr_threshold = 0.05)
}

```

Find *knee point* and *inflection* using BarcodeRanks

```{R}
path_list <- c("C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_noPEG_R1/Solo.out/GeneFull/raw",
               "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_noPEG_R2/Solo.out/GeneFull/raw",
                "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_PEG_R1/Solo.out/GeneFull/raw",
                "C:/Maria/IB/metrics/3.STAR/Fixed_fresh_K562_PEG_R2/Solo.out/GeneFull/raw",
                "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_noPEG_R1/Solo.out/GeneFull/raw",
                "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_noPEG_R2/Solo.out/GeneFull/raw",
                "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_PEG_R1/Solo.out/GeneFull/raw",
                "C:/Maria/IB/metrics/3.STAR/Fixed_frozen_K562_PEG_R2/Solo.out/GeneFull/raw")

inflection_points <- c()
knee_points <- c()

for (i in 1:length(path_list)) {
  # Read 10x data
  sce <- read10xCounts(path_list[i], type = "sparse")
  br.out <- barcodeRanks(assays(sce)$counts)
  inflection <- br.out@metadata$inflection
  knee <- br.out@metadata$knee
  inflection_points[i] <- inflection
  knee_points[i] <- knee
}
print("knee_points:")
print(knee_points)
print('inflection_points:')
print(inflection_points)
```

Data filtration with BR inflection point and knee_point
```{R, BR filtration}
BR_selection_inflection <- list()

for (i in 1:length(inflection_points)) {
  var_name <- paste0("inf_BR_s", i)
  BR_selection_inflection[[i]] <- subset(get(paste0("s", i)), get(paste0("s", i))$umi_per_CB > inflection_points[i])
  assign(var_name, BR_selection_inflection[[i]])
}

BR_selection_knee <- list()

for (i in 1:length(knee_points)) {
  var_name <- paste0("knee_BR_s", i)
  BR_selection_knee[[i]] <- subset(get(paste0("s", i)), get(paste0("s", i))$umi_per_CB > knee_points[i])
  assign(var_name, BR_selection_knee[[i]])
}

```


Selected barcodes in non filtered, > 100 UMI in each
```{R, >100UMI}

s_100 <- list()

for (i in 1:8) {
  var_name <- paste0("s", i, "_100")
  s_100[[i]] <- subset(get(paste0("s", i)), get(paste0("s", i))$umi_per_CB > 100)
  assign(var_name, s_100[[i]])
}

```

Calculate some sample characteristics to compare

```{R,  samples characteristics}

max_umi_count = c()
for (i in 1:length(s)) {
  smax <- max(s[[i]]$umi_per_CB)
  max_umi_count[i] <- smax
}

s_BR = list(inf_BR_s1, inf_BR_s2, inf_BR_s3, inf_BR_s4, inf_BR_s5, inf_BR_s6, inf_BR_s7, inf_BR_s8)
average_UMI_count_more_than_inflection = c()
for (i in 1:length(s_BR)) {
  smean <- mean(s_BR[[i]]$umi_per_CB)
  average_UMI_count_more_than_inflection[i] <- smean
}

average_UMI_count = c()
for (i in 1:length(s)) {
  smean <- mean(s[[i]]$umi_per_CB)
  average_UMI_count[i] <- smean
}

```

Calculation for number of barcodes selected by ED depending on UMI_per_CB counts:
in 3 groups:
100 < UMI_per_CB < inflection_point
inflection_point < UMI_per_CB < knee_point
knee_point < UMI_per_CB 


```{R}
count_between_knee_and_inflection <- c()
count_between_knee_and_inflection_ED_yes <- c()
count_between_knee_and_inflection_ED_no <- c()
count_more_than_knee <- c()
count_more_than_knee_ED_yes <- c()
count_more_than_knee_ED_no <- c()
count_between_100_and_inflection <- c()
count_between_100_and_inflection_ED_yes <- c()
count_between_100_and_inflection_ED_no <- c()

for (i in 1:8) {

  ED_si <- subset(s[[i]], row.names(s[[i]]) %in% ED_CB_list[[i]])
  si <- get(paste0("s", i))
  
  
  count_between_knee_and_inflection_ED_yes[i] <- dim(subset(ED_si, 
                                        inflection_points[i] < umi_per_CB & umi_per_CB < knee_points[i]))[1]
  count_between_knee_and_inflection[i] <- dim(subset(si, 
                                        inflection_points[i] < umi_per_CB & umi_per_CB < knee_points[i]))[1]
  count_between_knee_and_inflection_ED_no[i] <- count_between_knee_and_inflection[i] -     count_between_knee_and_inflection_ED_yes[i]
  
  
  
  count_more_than_knee_ED_yes[i] <- dim(subset(ED_si, 
                                        knee_points[i] < umi_per_CB))[1]
  count_more_than_knee[i] <- dim(subset(si, 
                                        knee_points[i] < umi_per_CB))[1]
  count_more_than_knee_ED_no[i] <- count_more_than_knee[i] - count_more_than_knee_ED_yes[i]
  
  
  
  count_between_100_and_inflection_ED_yes[i] <- dim(subset(ED_si, 
                                        100 < umi_per_CB & umi_per_CB < inflection_points[i]))[1]
  count_between_100_and_inflection[i] <- dim(subset(si, 
                                        100 < umi_per_CB & umi_per_CB < inflection_points[i]))[1]
  count_between_100_and_inflection_ED_no[i] <- count_between_100_and_inflection[i] -     count_between_100_and_inflection_ED_yes[i]
}
print(count_between_knee_and_inflection)
print(count_between_knee_and_inflection_ED_yes)
print(count_between_knee_and_inflection_ED_no)

print(count_more_than_knee)
print(count_more_than_knee_ED_yes)
print(count_more_than_knee_ED_no)

print(count_between_100_and_inflection)
print(count_between_100_and_inflection_ED_yes)
print(count_between_100_and_inflection_ED_no)

```


Now We are compiling all the collected data into a csv file for further analysis and visualization of the disproportionate selection of non-empty droplets in different samples by the Empty Drops package.
```{R}
sum_data <- data.frame(
    "sample_name" = c(1, 2, 3, 4, 5, 6, 7, 8),
    "knee_point" = knee_points,
    "inflection_point" = inflection_points,
    "max_umi_per_CB_count" = max_umi_count,
    "average_UMI_count_more_than_inflection" = average_UMI_count_more_than_inflection,
    "average_UMI_count" = average_UMI_count,
    "count_between_knee_and_inflection" = count_between_knee_and_inflection,
    "count_between_knee_and_inflection_ED_yes" = count_between_knee_and_inflection_ED_yes,
    "count_between_knee_and_inflection_ED_no" = count_between_knee_and_inflection_ED_no,
    "count_more_than_knee" = count_more_than_knee,
    "count_between_100_and_inflection" = count_between_100_and_inflection,
    "count_between_100_and_inflection_ED_yes" = count_between_100_and_inflection_ED_yes,
    "count_between_100_and_inflection_ED_no" = count_between_100_and_inflection_ED_no)

sum_data %>% write.csv("8_samples_summarized_data.csv")
```
