---
title: "PDF_distance_graphs"
author: "Shakir Suleimanov"
date: "2024-05-08"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```


# Import libraries
```{r libraries}
library(ggplot2)
library(readr)
library(dplyr)
```


```{r}
inflection_points <- c(843, 512, 409, 395, 1308, 1143, 325, 231, 646, 325)

knee_points <- c(4898, 3265, 921, 798, 2287, 2401, 511, 439, 1601, 1513)
```


# Import datasets for correlation distances
```{r}
ED_logs_Solo_out_1 <- read.csv("../distances/3000hvg/data_with_LogProbs_ED_between_inflection_and_knee_points_3000hvg_Solo.out_1.csv")

ED_logs_Solo_out_2 <- read.csv("../distances/3000hvg/data_with_LogProbs_ED_between_inflection_and_knee_points_3000hvg_Solo.out_2.csv")

ED_logs_Fixed_fresh_K562_noPEG_R1 <- read.csv("../distances/3000hvg/data_with_LogProbs_ED_between_inflection_and_knee_points_3000hvg_Fixed_fresh_K562_noPEG_R1.csv")

ED_logs_Fixed_fresh_K562_noPEG_R2 <- read.csv("../distances/3000hvg/data_with_LogProbs_ED_between_inflection_and_knee_points_3000hvg_Fixed_fresh_K562_noPEG_R2.csv")

ED_logs_Fixed_fresh_K562_PEG_R1 <- read.csv("../distances/3000hvg/data_with_LogProbs_ED_between_inflection_and_knee_points_3000hvg_Fixed_fresh_K562_PEG_R1.csv")

ED_logs_Fixed_fresh_K562_PEG_R2 <- read.csv("../distances/3000hvg/data_with_LogProbs_ED_between_inflection_and_knee_points_3000hvg_Fixed_fresh_K562_PEG_R2.csv")

ED_logs_Fixed_frozen_K562_noPEG_R1 <- read.csv("../distances/3000hvg/data_with_LogProbs_ED_between_inflection_and_knee_points_3000hvg_Fixed_frozen_K562_noPEG_R1.csv")

ED_logs_Fixed_frozen_K562_noPEG_R2 <- read.csv("../distances/3000hvg/data_with_LogProbs_ED_between_inflection_and_knee_points_3000hvg_Fixed_frozen_K562_noPEG_R2.csv")

ED_logs_Fixed_frozen_K562_PEG_R1 <- read.csv("../distances/3000hvg/data_with_LogProbs_ED_between_inflection_and_knee_points_3000hvg_Fixed_frozen_K562_PEG_R1.csv")

ED_logs_Fixed_frozen_K562_PEG_R2 <- read.csv("../distances/3000hvg/data_with_LogProbs_ED_between_inflection_and_knee_points_3000hvg_Fixed_frozen_K562_PEG_R2.csv")
```


# Create full DF for eucl distances for 2 samples
```{r}
ED_logs_dataset_2s <- rbind(
  ED_logs_Solo_out_1,
  ED_logs_Solo_out_2
)
```


# Create full DF for cor distances for 8 samples
```{r}
ED_logs_dataset_8s <- rbind(
  ED_logs_Fixed_fresh_K562_noPEG_R1,
  ED_logs_Fixed_fresh_K562_noPEG_R2,
  ED_logs_Fixed_fresh_K562_PEG_R1,
  ED_logs_Fixed_fresh_K562_PEG_R2,
  ED_logs_Fixed_frozen_K562_noPEG_R1,
  ED_logs_Fixed_frozen_K562_noPEG_R2,
  ED_logs_Fixed_frozen_K562_PEG_R1,
  ED_logs_Fixed_frozen_K562_PEG_R2
)
```


# Create density plot for correlation distances for 2 samples
```{r fig.height=6, fig.width=10}
ggplot(ED_logs_dataset_2s, aes(x = Distance)) +
  geom_density(aes(fill = as.factor(Sample)), alpha = 0.7) +
  theme_bw() +
  labs(fill = "Samples") +
  xlab("LogProbabilities")
```



# Create density plot for correlation distances for 8 samples
```{r fig.height=6, fig.width=10}
ggplot(ED_logs_dataset_8s, aes(x = Distance)) +
  geom_density(aes(fill = as.factor(Sample)), alpha = 0.7) +
  theme_bw() +
  labs(fill = "Samples") +
  xlab("LogProbabilities")
```


# Scatter plot for correlation distance for 2 samples
```{r fig.height=6, fig.width=10}
ggplot(ED_logs_dataset_2s, aes(y = Distance, x = UMI_count, fill = Sample)) +
  geom_point(aes(color = Sample), alpha = 0.7) +
  geom_smooth(method = "lm") +
  theme_bw() +
  xlab("LogProbabilities")
```



# Scatter plot for correlation distance for 8 samples
```{r fig.height=6, fig.width=10}
ggplot(ED_logs_dataset_8s, aes(y = Distance, x = UMI_count, fill = Sample)) +
  # geom_point(aes(color = Sample), alpha=0.4) +
  geom_smooth(method = "lm") +
  theme_bw() +
  xlab("LogProbabilities")
```



