---
title: " Knee-plots coloured by UMI distribution between 2 samples"
author: "Maria Lukina"
date: "2024-04-25"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Libraries and Imports

```{r, libraries, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(tibble)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(viridis)
library(RColorBrewer)
library(glue)
source('../R/read_data.R')
source('../R/emptyDrops_filtering.R')
```

We are going to analyse EmptyDrops working on 2 samples: Solo1 and Solo2
##### Organizing data

```{R}
#-----------------------------------------------------Solo1_raw

counts_path_s1 <- ("../Data/Solo.out_1/Gene/raw/matrix.mtx.gz")
genes_path_s1 <- ("../Data/Solo.out_1/Gene/raw/features.tsv.gz")
barcodes_path_s1 <- ("../Data/Solo.out_1/Gene/raw/barcodes.tsv.gz")

#-----------------------------------------------------Solo2_raw
counts_path_s2 <- ("../Data/Solo.out_2/Gene/raw/matrix.mtx.gz")
genes_path_s2 <- ("../Data/Solo.out_2/Gene/raw/features.tsv.gz")
barcodes_path_s2 <- ("../Data/Solo.out_2/Gene/raw/barcodes.tsv.gz")

```

Data reading  and EmptyDrops running
```{R matrices,  warning=FALSE, , message=FALSE}
#Solo1
matrix_1_raw <- sc_matrix_10X(counts_path_s1, barcodes_path_s1, genes_path_s1)
emptyDrops_df_1 <- filter_barcodes_with_emptyDrops(matrix_1_raw)
filtered_barcodes1 <- return_filtered_barcodes_or_indices(emptyDrops_df_1, fdr_threshold = 0.05) 
#Solo2
matrix_2_raw <- sc_matrix_10X(counts_path_s2, barcodes_path_s2, genes_path_s2)
emptyDrops_df_2 <- filter_barcodes_with_emptyDrops(matrix_2_raw)
filtered_barcodes2 <- return_filtered_barcodes_or_indices(emptyDrops_df_2, fdr_threshold = 0.05) 

```
Counting the total number of UMI for each barcode and ranking the barcodes for each sample

```{R, UMI_per_CB count}

#Calculate UMI per cell barcode distribution
umi_per_CB_1 <- calculate_umi_per_cb(matrix_1_raw)
umi_per_CB_2 <- calculate_umi_per_cb(matrix_2_raw)

#Create sorted UMI_count DF
DF_umi_per_CB1 <- umi_per_CB_1 %>%
  arrange(desc(umi_per_CB_1)) %>%
  mutate(cb = as.numeric(seq_len(nrow(umi_per_CB_1))))

DF_umi_per_CB2 <- umi_per_CB_2 %>%
  arrange(desc(umi_per_CB_2)) %>%
  mutate(cb = as.numeric(seq_len(nrow(umi_per_CB_2))))

```
Creating a combined dataframe where barcodes are numbered according to the total number of UMI from both samples, and the proportion of UMI coming from each sample is calculated.

```{R, combine 2 samples}
DF_umi_per_CB_s1_s2_sort <- merge(DF_umi_per_CB1, DF_umi_per_CB2, 
                             by = 'row.names', all = TRUE)

#Replcase NA's with 0
DF_umi_per_CB_s1_s2_sort[is.na(DF_umi_per_CB_s1_s2_sort)] <- 0

#Calculate distribution and range Barcodes in summarized sampe

DF_all <- DF_umi_per_CB_s1_s2_sort %>%
  mutate(dist_s1 = round(UMIcount.x/(UMIcount.x + UMIcount.y),6)) %>%
  mutate(dist_s2 = round(UMIcount.y/(UMIcount.x + UMIcount.y),6))%>%
  mutate(umi_per_CB_total = UMIcount.y + UMIcount.x)
DF_all_sorted <- as.data.frame(DF_all) %>%
  arrange(desc(DF_all$umi_per_CB_total))
DF_final <- as.data.frame(DF_all_sorted) %>%
  mutate(common_CB_number = as.numeric(rownames(as.data.frame(DF_all_sorted))))
```

 
####Create knee-plots coloured by proportion of UMIs coming from sample Solo1 for selected/not selected by ED and all barcodes

Now, let's define a function that  build a knee_plot and color it depending on the proportion of UMI coming from the first sample in the "total sample".

The function will take a dataframe, maximum value for y scale, and graph title as input.

The df should have the following columns:
* `common_CB_number`: the barcode number from the combined df for the two samples, sorted by the total number of UMI/CB
* `umi_per_CB`: the number of UMI for this barcode in combined sample
* `dist_s1`: the ratio of the number of UMI for this barcode in sample 1 to the total number of UMI for this barcode
* `y_max`: the maximum value for y scale
```{R}
knee_plot_UMI_sample_dist <- function(df, y_max, name){
  colors <- brewer.pal(n = 11, name = "Spectral")
  knee_plot_UMI_sample_dist <- ggplot(df, aes(x = common_CB_number, y = umi_per_CB_total, colour = dist_s1*100)) +
    geom_point(size = 2) +
    labs(y = 'UMI count, log10', x = 'cell barcodes, log10', title = name) +
    theme_linedraw() +
    theme(axis.text=element_text(size=14),
          axis.title=element_text(size=16,face="bold")) +
    scale_y_log10(limits = c(100, y_max)) +
    scale_x_log10(limits = c(1, NA)) +
    scale_colour_gradientn(colours = colors,
                           guide = guide_colorbar(title = "UMIs from Sample1, %")) 
  return (knee_plot_UMI_sample_dist)
}
```
Preparing dataframes for 'summarized sample' containing > 100 UMIs (all_100) and its to subsets: barcodes selected by ED (yes) and not selected (no)

```{R, message = FALSE, warnings = FALSE}
filtered_barcodes <- unique(c(filtered_barcodes1, filtered_barcodes2))
ED_yes <-  subset(DF_final, Row.names %in% filtered_barcodes)
all_100 <-  subset(DF_final, DF_final$UMIcount.x > 100 | DF_final$UMIcount.y > 100)
ED_no <-  subset(all_100, !(Row.names %in% filtered_barcodes))
# mixed to avoid trends in point overlapping on plot
ED_yes <- ED_yes[sample(nrow(ED_yes)), ]
ED_no <- ED_no[sample(nrow(ED_no)), ]
all_100 <- all_100[sample(nrow(all_100)), ]

```
Visualization by knee_plots
For better visualization, we will plot all three graphs in the same interval on the y-axis: from 100 to the maximum number of UMI in the combined sample
```{R, knee_plots, warning = FALSE}
y_max <-  max(all_100$umi_per_CB_total)+ 200
yes <- knee_plot_UMI_sample_dist(ED_yes, y_max, 'Selected by ED, (ED_yes)') 

no <-  knee_plot_UMI_sample_dist(ED_no, y_max, 'Not selected by ED (ED_no)') 

all <- knee_plot_UMI_sample_dist(all_100, y_max, 'all, nUMIs > 100') 

yes
no
all

combined_plot_3 <- (yes |  no | all) + 
  plot_annotation(
    theme = theme(plot.title = element_text(hjust = 0.5)) 
  )


file_path <- "knee_plots_all_yes_no.png"
ggsave(file = file_path, plot = combined_plot_3, width = 24, height = 7, units = "in", dpi = 300)
```
