---
title: "Distances between EmptyDrops and non_empty drops in sample 1"
author: "Shakir Suleimanov"
date: "2024-03-26"
output: 
  html_document:
    #css: styles.css
    latex_engine : xelatex
    df_print: default
    highlight: zenburn
    toc: TRUE
    toc_depth: 4
mainfont : NanumGothic    
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Libs
```{r message=FALSE, warning=FALSE, include=FALSE}
source('../R/calculate_distances.R')
```


### Datapaths
```{r paths, echo=TRUE, message=FALSE, warning=FALSE}
data <- read.csv('../distances/3000hvg/data_with_cor_dist_between_inflection_and_knee_points_3000hvg_Fixed_fresh_K562_noPEG_R1.csv', row.names = 1)

```


### Get barcodes which pass the EmptyDrops filtering
```{r}
empty_drops_df_sample_1 <- filter_barcodes_with_emptyDrops(sc_sample_1)

not_empty_barcodes_sample_1 <- return_filtered_barcodes_or_indices(empty_drops_df_sample_1)
```




# Plot knee_plot for cells with UM_count over 100 - correlation distance
```{r fig.height=10, fig.width=10}
knee_plot_with_distances(data_with_cor_dist_for_plot_sc_sample_over_100, not_empty_barcodes_sample)
```


# Plot knee_plot for cells with UM_count over 1000 - correlation distance
```{r fig.height=10, fig.width=10}
knee_plot_with_distances(data_with_cor_dist_for_plot_sc_sample_over_1000, not_empty_barcodes_sample)
```


# Plot knee_plot for cells with UM_count over 3000 - correlation distance
```{r fig.height=10, fig.width=10}
knee_plot_with_distances(data_with_cor_dist_for_plot_sc_sample_over_3000, not_empty_barcodes_sample)
```


# Plot knee_plot for cells with UM_count over inflection point - correlation distance
```{r fig.height=10, fig.width=10}
knee_plot_with_distances(data_with_cor_dist_for_plot_sc_sample_over_ip, not_empty_barcodes_sample)
```


# Calculate correlation coefficients between UMI_counts and distance (correlation) for dataset with cell (over 100, over 1000 and over 3000)
```{r}
cor(data_with_cor_dist_for_plot_sc_sample_over_100$UMI_count, data_with_cor_dist_for_plot_sc_sample_over_100$Distance, method = 'spearman', use='pairwise.complete.obs')

cor(data_with_cor_dist_for_plot_sc_sample_over_1000$UMI_count, data_with_cor_dist_for_plot_sc_sample_over_1000$Distance, method = 'spearman')

cor(data_with_cor_dist_for_plot_sc_sample_over_3000$UMI_count, data_with_cor_dist_for_plot_sc_sample_over_3000$Distance, method = 'spearman')

cor(data_with_cor_dist_for_plot_sc_sample_over_ip$UMI_count, data_with_cor_dist_for_plot_sc_sample_over_ip$Distance, method = 'spearman')
```


# Plot knee_plot for cells with UM_count over 100 - normalized euclidean distance
```{r fig.height=10, fig.width=10}
knee_plot_with_distances(data_with_eucl_dist_for_plot_sc_sample_over_100, not_empty_barcodes_sample)
```


# Plot knee_plot for cells with UM_count over 1000 - normalized euclidean distance
```{r fig.height=10, fig.width=10}
knee_plot_with_distances(data_with_eucl_dist_for_plot_sc_sample_over_1000, not_empty_barcodes_sample)
```


# Plot knee_plot for cells with UM_count over 3000 - normalized euclidean distance
```{r fig.height=10, fig.width=10}
knee_plot_with_distances(data_with_eucl_dist_for_plot_sc_sample_over_3000, not_empty_barcodes_sample)
```


# Plot knee_plot for cells with UM_count over inflection point - normalized euclidean distance
```{r fig.height=10, fig.width=10}
knee_plot_with_distances(data_with_eucl_dist_for_plot_sc_sample_over_ip, not_empty_barcodes_sample)
```


# Calculate correlation coefficients between UMI_counts and distance (euclidean) for dataset with cell (over 100, over 1000 and over 3000)
```{r}
cor(data_with_eucl_dist_for_plot_sc_sample_over_100$UMI_count, data_with_eucl_dist_for_plot_sc_sample_over_100$Distance, method = 'spearman', use='pairwise.complete.obs')

cor(data_with_eucl_dist_for_plot_sc_sample_over_1000$UMI_count, data_with_eucl_dist_for_plot_sc_sample_over_1000$Distance, method = 'spearman')

cor(data_with_eucl_dist_for_plot_sc_sample_over_3000$UMI_count, data_with_eucl_dist_for_plot_sc_sample_over_3000$Distance, method = 'spearman')

cor(data_with_eucl_dist_for_plot_sc_sample_over_ip$UMI_count, data_with_eucl_dist_for_plot_sc_sample_over_ip$Distance, method = 'spearman')
```










